---
title: "Nationwide NEON Mosquito Annual Trends: 2016 - 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Load packages
library(tidyverse)
library(lubridate)
library(janitor)
library(plotly)
library(viridis)
library(patchwork)
library(scales)


## Working directory for .RMD
knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#Set Scientific notation output and decimal places for knitr
options(scipen = 999)
options(digits = 4)
options(dplyr.summarise.inform = FALSE)

```

```{r data_read, include = FALSE}

## Read in mosquito data list
mosq_data <- read_rds("./data/mosq_data.rds")

## Grab the sorting data
sort_data <- tibble(mosq_data$mos_sorting) %>%
  janitor::clean_names() %>%
  mutate(
    year = year(collect_date), ## Grab year of collection
    biomass = (total_weight - bycatch_weight) * 1000, ## Get mosquito mass in mg
    trap_hours = as.numeric(collect_date - set_date), ## Get trapping time
    biomass_per_hour = biomass / trap_hours ## Standardize biomass by trapping time
  ) %>%
  filter(
    trap_hours != 0 ## Remove erroneous entries with absent trapping times
  )

## Grab the ID data
id_data <- tibble(mosq_data$mos_expertTaxonomistIDProcessed) %>%
  janitor::clean_names() %>%
  mutate(
    year = year(collect_date)
  )
```

```{r data_process, include = FALSE}

## Summarize biomass by year
biomass_summary <- sort_data %>%
  group_by(year) %>%
  summarize(
    total_bm = sum(biomass),
    total_bm_kg = total_bm / 1000000,
    total_hrs = sum(trap_hours),
    bm_per_hr = total_bm / total_hrs
  ) %>%
  ungroup()


## Summarize mosquito counts by year
count_summary <- id_data %>%
  filter(!is.na(genus)) %>%
  mutate(
    genus = genus %>% factor() %>% fct_collapse(Other = c("Mansonia", "Orthopodomyia", "Uranotaenia", "Wyeomyia"))
  ) %>%
  group_by(year, genus) %>%
  summarize(
    individual_count = sum(individual_count)
  ) %>%
  mutate(
    genus = genus %>% factor() %>% fct_reorder(.x = individual_count, .fun = sum, .desc = TRUE),
    text_label = str_c("Genus: ", genus,
                            "\n# IDed: ", formatC(individual_count, big.mark = ",", format = "fg"),
                            "\nYear: ", year)
    )

```

### Annual Trends in Biomass

```{r bm_plots, echo = FALSE}

## Plot total biomass
bm_tot <- biomass_summary %>%
  ggplot() +
  geom_col(aes(x = year, y = total_bm_kg, fill = year)) +
  scale_fill_viridis(option = "cividis") +
  labs(x = "Year", y = "Total Biomass (kg)") +
  coord_flip(ylim = c(max(biomass_summary$total_bm_kg) * 1.1, 0),
             xlim = c(max(biomass_summary$year) + 0.7, min(biomass_summary$year) - 0.7), 
             expand = FALSE) +
  scale_x_reverse() +
  scale_y_reverse() +
  theme_bw() +
  theme(legend.position = "none")

## Plot biomass by hour
bm_hr <- biomass_summary %>%
  ggplot() +
  geom_col(aes(x = year, y = bm_per_hr, fill = year)) +
  scale_fill_viridis(option = "cividis") +
  labs(x = "Year", y = "Total Biomass by Trapping Time (mg/hr)") +
  coord_flip(ylim = c(0, max(biomass_summary$bm_per_hr) * 1.1),
             xlim = c(max(biomass_summary$year) + 0.7, min(biomass_summary$year) - 0.7), 
             expand = FALSE) +
  theme_bw() +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(hjust = -0.01))

## Combine plots
bm_tot + bm_hr

```

Total biomass is calculated as the sum of mosquito biomass collected in all traps across all NEON domains for each year (in kilograms). Total biomass by trapping time is that value (in milligrams) divided by the time spent trapping across all domains. 

</br>

### Annual Trends in Number of Individuals Identified by Genus

```{r species_plot, echo = FALSE, warning = FALSE}

## Plot number mosquitoes IDed by genus
count_plot <- count_summary %>%
  ggplot() +
  geom_col(aes(x = year, y = individual_count, fill = genus, text = text_label)) +
  scale_fill_viridis_d(option = "plasma") + 
  scale_y_continuous(labels = comma) +
  labs(x = "Year", y = "# Individuals Identified") +
  theme_bw()

## Make plotly
ggplotly(count_plot, tooltip = "text") %>%
  layout(legend = list(title = list(text = "Genus")))

```

Note: "Other" genera include Mansonia, Orthopodomyia, Uranotaenia, and Wyeomyia.